  ---
title: "Practicals 1"
author: "Pietro Franceschi"
date: "29/09/2019"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The objective here is to start familaiarizing with R, and Rstudio As you see the characetr '#' is udes to idetify lines of comments they are not interpreted and are useful for us to kee track of the process of analysis

As discussed assignment (sending somethingin the workspace) is done using '<-' the shrotcut to insert it is 'ALT' + '-' 
a complete list of shortcut can be found in the menu under 'tools'

```{r}
## BASIC ASSIGNMENTS  #########################
a <- 1
A <- 2
mystring <- "fmach"   ## note the double quotes
myboolean <- TRUE
```

run the code and ispect your workspace (environnment) as you see R is case sensitive ...


## Multidimensional Objects
A simple multidimensional object (a vector) is created by a function c()

```{r}
myvector <- c(1,2,3,4)

## and this is the help for c!
?c

```


As we discussed, **multidimensional** objects can be accessed either by position of by name. This second way is possible only if the objects inside the multidimensionale "container" have names ...
```{r}
## assignment by name
u <- c("carlo" = 2, "due" = 10)

## how can I look inside a multidimensional object?
myvector[3]

## or only change the second element of my object
myvector[2] <- 30

## this gives me the third element of the  vector, remember that the first element in the 
## remember that R is counting from one and not zero
## if I want to construct a 2d array ...

myarray <- matrix(seq(1,25,1), nrow = 5, ncol = 5)

myarray

```


Now I would like you to comment the previous command .....

Key ideas:
- what is a function
- function arguments
- function default arguments


More on multidimensional objects ...

```{r}
## the content can be accessed
myarray[2,3] ## one element
myarray[1,]  ## one row!

## square brakets can be used in a quite flexible way to select elements
## matching specific properties

myarray[myarray > 10]
myarray[myarray[1,] >3,]

## when you have multidimensional objects it is often useful give
## a name to the eleents

vector.with.names <- c("one" = 1, "two" = 2, "three" = 3)

## in this way I can access the second element either as
vector.with.names[2]
##or
vector.with.names["two"]

## with two dimensional or more dimensional objects rownames and colum names are useful
## exactly as in excel
```



## The first assignment!

```{r}

mymonths <- c("Jan","Feb","Mar",
              "Apr","May","Jun",
              "Jul","Aug","Sep",
              "Oct","Nov","Dec")
days <- c(31,28,31,30,31,30,31,31,30,31,30,31)
season <- c(rep(c("winter","spring","summer","autumn"), each = 3))

yeardf <- data.frame("month" = mymonths, "length" = days, "season" = season,
                     stringsAsFactors = FALSE)

factor(yeardf$season)


## if I want to use a list ...
yearlist <- list("month" = mymonths, "length" = days, "season" = season)
yearlist[["length"]]

```


## Data visualization

```{r}
## I load my first package!
library(tidyverse)

## and also some data
data(mpg)

## as we discussed this shows the first lines of the data.frame
head(mpg)
```

## Description
This dataset contains a subset of the fuel economy data that the EPA makes available on http://fueleconomy.gov. It contains only models which had a new release every year between 1999 and 2008 - this was used as a proxy for the popularity of the car.


```{r}
## a first glance on the data
summary(mpg)
```

## Plots!
And now we look at plots!

```{r}
## my first plot .... engine displacement, miles per gallon
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), col = "red")
  
```



This goes a little bit farer away

```{r}
ggplot(data = mpg) +
  ## on the canvas plot points linking their eastetics to the column names!
  geom_point(mapping = aes(x = displ, y = hwy, col = class),  
             size = 3, 
             alpha = 0.5) +
  ## add additional general aestetics
  theme_light() +
  theme(aspect.ratio = 0.5)

```



what aesthetics could I use? another time ....

?geom_point


```{r}
## histograms, scales free
ggplot(data = mpg) +
  geom_histogram(mapping = aes(x = displ, fill = class), alpha = 0.5) 
```


```{r}
## or jitter!
ggplot(data = mpg) +
  geom_jitter(mapping = aes(x = class, y = displ, col = class), width = 0.1)

```


This is some "material" for your assignment ;-)

```{r}
## geom_jitter
## geom_histogram
## geom_line
## geom_point
## geom_density
```


## Assignment #1

1. Play around with mpg changing the type of plots 
2. Associate aesthetics to categorical or continuous properties
3. Make a boxplot of "displ" as a function of the class of the vehicle
4. Just play and ask!


## Splitting the plot in subplots

As we discussed, faceting is used to split a plot in different subpanels on the bases of a categorical variable, which is normally present in the data.frame/tibble

```{r}
## group specific subplots
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, col = class), alpha = 0.5) +
  facet_wrap(~class) + 
  theme_light() +
  theme(aspect.ratio = 0.5)
```

With the previous command all the plots have the same scale, this behavior can be controlled by setting the `scales` argument inside the facet_wrap() call.

## From wide to long data.frames

As we discussed in the lecture, a clever combination of reshaping and faceting can be used to plot in a unique plot a series of different properties (column).

To do that, a data frame in wide form should be transformed in its "long" counterpart.

```{r}
## the function "gather" goes from wide to long format. To do that it is necessary to provide the name 
## of two new columns: the one which will store the variable names and the one which will collect the values
## here we call them "Property" and "value"
mpglong <- gather(mpg, 
                  key = "Property", 
                  value = "Value",
                  displ,cty,hwy)  ## these are the names of the column we are "collecting"

mpglong
```


## Make a full plot of all the properties ...

We use now the long data frame 

```{r fig.width=12}
ggplot(mpglong) + 
  geom_jitter(aes(x = class, y = Value, col = class), width = 0.1) + 
  geom_boxplot(aes(x = class, y = Value, fill = class), alpha = 0.5) + 
  facet_wrap(~Property, scales = "free") + 
  coord_flip() +         ### switch the axis to show the names ...
  theme_light() + 
  theme(aspect.ratio = 1)
```


## Iris data
The last part of this pratical is dedicated to the Iris dataset, which is also available in R

```{r}
## here is the starting of the story ....
data(iris)

head(iris)
```

## Assignment #4
1. Look for correlation between  Sepal.Length and Sepal.Width (Petal.Length and Petal.Width) for the three iris varieties
2. Make a faceted boxplot showing the four iris properties as a function of the species  






