---
title: "practical_3"
author: "Pietro Franceschi"
date: "03/10/2019"
output: html_document
---

## Loads

```{r}
## libraries
library(tidyverse)
library(readr)
```


```{r}
## Olympic games data
athl <- read_csv("data/athlete_events.csv")
```

How many specialities so we have?

```{r}
## number of specialities
length(unique(athl$Event))

## number of athlets
nrow(athl)
```

In athletics, can we see a change in the age of the athlets over time?

```{r}
## let's get the list of the sports
unique(athl$Sport)
```


```{r}
athl %>% 
  filter(Sport == "Athletics") %>% 
  ggplot() + 
  geom_jitter(aes(x = Year, y = Age, col = Sex), width = 0.2, size = 0.1) +
  scale_color_brewer(palette = "Set1", name = "Gender") + 
  facet_wrap(~Sex, ncol = 1) + 
  theme_light()
  
```

Which, by the way, gives already some interesting result ... apparently the age of participation is rising mainly in females


```{r}
athl %>% 
  filter(Sport == "Athletics") %>% 
  filter(Medal %in% c("Gold","Silver","Bronze")) %>% 
  ggplot() + 
  geom_jitter(aes(x = Year, y = Age, col = Sex), width = 0.2, size = 0.1) +
  geom_smooth(aes(x = Year, y = Age), method = lm) + 
  scale_color_brewer(palette = "Set1", name = "Gender") + 
  facet_wrap(~Sex, ncol = 1) + 
  theme_light()
```

## Grouping and Summarising

As an easy starting point we start from the iris dataset

```{r}
## get the data
data(iris)
```

Now we group it ...

```{r}
iris %>% 
  group_by(Species)
```

Apparently nothing has changed ...

```{r}
## This performs "on the fly" the mean per group! 

iris %>% 
  group_by(Species) %>% 
  summarise(sep = mean(Sepal.Length))

## remember if we not assign the pipe to something the output will be only printed!
```


```{r}
## This performs "on the fly" the mean and sample counts per group! 
iris %>% 
  group_by(Species) %>% 
  summarise(sep = mean(Sepal.Length), 
            nsamp = length(Sepal.Length))
```

Multiple columns can be summarised by using `summarise_at`

```{r}
iris %>% 
  group_by(Species) %>% 
  summarise_at(vars(`Sepal.Length`:`Petal.Width`),
               .funs = list(sep = mean, 
                            nsamp = length))
```

This type of writing ca be challenging ... the alternative is to rely on gather ;-)


```{r}
iris %>% 
  gather(parameter,value, -Species) %>% 
  group_by(parameter,Species) %>% 
  summarise(sep = mean(value), 
            nsamp = length(value)) %>% 
  
```


```{r}
athl %>% 
  filter(Season == "Summer") %>% 
  filter(Medal %in% c("Gold")) %>% 
  group_by(Year,Name,Sport) %>% 
  summarise(medals = length(Medal)) %>% 
  ungroup() %>% 
  filter(medals > 3) %>% 
  arrange(desc(medals))
```



