---
title: "Practicals 1"
author: "Pietro Franceschi"
date: "29/09/2019"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The objective here is to start familaiarizing with R, and Rstudio As you see the characetr '#' is udes to idetify lines of comments
they are not interpreted and are useful for us to kee track of the process  of analysis

As discussed assignment (sending somethingin the workspace) is done using '<-' the shrotcut to insert it is 'ALT' + '-' 
a complete list of shortcut can be found in the menu under 'tools'

```{r}
## BASIC ASSIGNMENTS  #########################
a <- 1
A <- 2
mystring <- "fmach"   ## note the double quotes
myboolean <- TRUE
```

run the code and ispect your workspace (environnment) as you see R is case sensitive ...


## Multidimensional Objects
A simple multidimensional object (a vector) is created by a function c()

```{r}
myvector <- c(1,2,3,4)

## and this is the help for c!
?c

```


As we discussed, multidimensional objects can be accessed either by position of by name. This second way is possible only if the objects inside the multidimensionale "container" have names ...
```{r}
## assignment by name
u <- c("carlo" = 2, "due" = 10)

## how can I look inside a multidimensional object?
myvector[3]

## or only change the second element of my object
myvector[2] <- 30

## this gives me the third element of the  vector, remember that the first element in the 
## remember that R is counting from one and not zero
## if I want to construct a 2d array ...

myarray <- matrix(seq(1,25,1), nrow = 5, ncol = 5)

myarray

```


now I would like you to comment the previous command .....

Key ideas:
- what is a function
- function arguments
- function default arguments


More on multidimensional objects ...

```{r}
## the content can be accessed
myarray[2,3] ## one element
myarray[1,]  ## one row!

## square brakets can be used in a quite flexible way to select elements
## matching specific properties

myarray[myarray > 10]
myarray[myarray[1,] >3,]

## when you have multidimensional objects it is often useful give
## a name to the eleents

vector.with.names <- c("one" = 1, "two" = 2, "three" = 3)

## in this way I can access the second element either as
vector.with.names[2]
##or
vector.with.names["two"]

## with two dimensional or more dimensional objects rownames and colum names are useful
## exactly as in excel
```



## The first assignment!

```{r}

mymonths <- c("Jan","Feb","Mar",
              "Apr","May","Jun",
              "Jul","Aug","Sep",
              "Oct","Nov","Dec")
days <- c(31,28,31,30,31,30,31,31,30,31,30,31)
season <- c(rep(c("winter","spring","summer","autumn"), each = 3))

yeardf <- data.frame("month" = mymonths, "length" = days, "season" = season,
                     stringsAsFactors = FALSE)

factor(yeardf$season)


## if I want to use a list ...
yearlist <- list("month" = mymonths, "length" = days, "season" = season)
yearlist[["length"]]

```


## Data visualization

```{r}
## I load my first package!
library(tidyverse)

## and also some data
data(mpg)


head(mpg)
```

## Description
This dataset contains a subset of the fuel economy data that the EPA makes available on http://fueleconomy.gov. It contains only models which had a new release every year between 1999 and 2008 - this was used as a proxy for the popularity of the car.


```{r}
## a first glance on the data
summary(mpg)
```

## Plots!
And now we look at plots!

```{r}
## my first plot .... engine displacement, miles per gallon
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = hwy, col = year), size = 3, alpha = 0.5) +
  theme_light()

```


```{r}
## the same can be extended ... playing with arguments
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), col = "red")
  
```

```{r}
## but more interestingly ...
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, col = class)) 
```


what aesthetics could I use? another time ....

?geom_point



```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, col = class)) + 
  facet_wrap(~class, scales ="free")

```


## More Plots

```{r}
## histograms, scales free
ggplot(data = mpg) +
  geom_histogram(mapping = aes(x = displ)) + 
  facet_wrap(~class)
```


```{r}
## or jitter!
ggplot(data = mpg) +
  geom_jitter(mapping = aes(x = class, y = displ, col = class), width = 0.1)

```

when you specify the names of the columns pay attention to the apex!

 "ciao"
 'ciao'
 `ciao`
 
 are not equivalent
 
 
 ## Get the data in 
 
```{r}
library(readr)
wines <- read_csv("wines.csv")
```
 
 
 And now ...
 
```{r}
summary(wines)
```
 
 
 
 



